---
title: "Answers to exercises \"Dependent data: Mixed effect models\" part 1"
output:
  github_document:
    toc: true
    toc_depth: 2
---

# Exercise 1

## 1a
```{r}
mm1 <- nlme::Milk[nlme::Milk$Time<=3,]
mm1 <- mm1[mm1$Time!=2,]
unpa <- t.test(protein~Time,paired = FALSE, var.equal = TRUE, data = mm1)
unpa
unpa$stderr
```

## 1b
```{r}
pa <- t.test(protein~Time, paired=TRUE, data = mm1)
pa
pa$stderr
```

## 1c
The standard error for the paired case is smaller because the observations on timepoint 1 and on timepoint 3 are correlated:
```{r}
t1 <- mm1$protein[mm1$Time==1]
t3 <- mm1$protein[mm1$Time==3]
cor.test(t1,t3)
```
So the correlation between protein on time 1 and time 3 is 0.47 with a 95% two sided confidence interval of (0.28;0.63)

## 1d
```{r}
library(lme4)
fit <- lmer(protein~factor(Time)+(1|Cow),data=mm1)
summary(fit)
```
## 1d

With the random intercept model the result:

               Estimate Std. Error t value  
               ...
factor(Time)3 -0.43089    0.04071  -10.58

is the same as with the paired t-test. (The difference in sign is because the random intercept model catculates the difference in means beyween time 3 and time 1, whereas the pairde t-test takes this diffrence between time 3 and time 1)

# Exercise 2

## 2a
```{r}
Od <- nlme::Orthodont
library(ggplot2)
library(lme4)
ggplot(Od,aes(x=age,y=distance,group=Subject))+geom_line(aes(color=Sex))
fit.od <- lmer(distance~factor(age)+factor(Sex)+factor(age):factor(Sex)+(1|Subject), REML=FALSE, data = Od)
```

## 2b
```{r}
Od <- data.frame(Od,res=residuals(fit.od))
ggplot(Od,aes(sample=res))+
  stat_qq()+stat_qq_line(color="red")+
  labs(y="Residuals",
       title="Normal Probability plot")+
  theme_bw()
```

## 2c
```{r}
drop1(fit.od)
```

Interaction can be legt out.
```{r}
fit.od2 <- lmer(distance~factor(age)+factor(Sex)+(1|Subject), REML=FALSE, data = Od)
drop1(fit.od2)
```

```{r}
summary(fit.od2,correlation=FALSE)
confint(fit.od2)
```

# Exercise 3

## 3a

```{r}
mm3 <- grouseticks
fit.gr <- lmer(TICKS~HEIGHT+factor(YEAR)+(1|BROOD),data = mm3)
```

## 3b

```{r}
drop1(fit.gr,test="Chisq")
```

## 3c

```{r}
mm3 <- data.frame(mm3,res=residuals(fit.gr))
ggplot(mm3,aes(sample=res))+
  stat_qq()+stat_qq_line(color="red")+
  labs(y="Residuals",
       title="Normal Probability plot")+
  theme_bw()
```

This model does not fit the data well. Since THICKS is a count variable, a poisson model migth do a better job.
