---
title: "Answers to exercizes \"count data: poison models\" "
output:
  github_document:
    toc: true
    toc_depth: 2
---

# Exercise 1

## 1a
```{r}
adress <- "https://raw.github.com/janvdbroek/Generalized-Linear-Models/master/cancer"
p1 <- read.table(adress,header=TRUE)
cancer.fit <- glm(cases~factor(age)+factor(city)+
                    offset(log(pop/1000)),family=poisson,data=p1)
```
## 1b
```{r}
p1 <- data.frame(p1,res=resid(cancer.fit),fv=fitted(cancer.fit))
library(ggplot2)
ggplot(p1,aes(x=fv,y=res))+geom_point(aes(color=factor(age)))
```

## 1c
```{r}
drop1(cancer.fit)
```

City can be left out:
```{r}
cancer.fit2 <- glm(cases~factor(age)+offset(log(pop/1000)),
                   family=poisson,data=p1)
drop1(cancer.fit2)
```
Age is needed in the model.

## 1d
```{r}
summary(cancer.fit2)
confint(cancer.fit2)
```

For example: the log ratio mean number of cancer cases per 1000 individualsfor age 5 vs age 1 is 1.8472 with a 95% profile log-likelihood confidence interval of (1.39;2.31)

For the ratio's:
```{r}
exp(coef(cancer.fit2))
exp(confint(cancer.fit2))
```
The mean number of cancer cases per 1000 individuals in agegroup 5 is 6.34 largers as compared to age group1.

# Exrcise 2

## 2a
```{r}
adress <- "https://raw.github.com/janvdbroek/Generalized-Linear-Models/master/coronary"
p2 <- read.table(adress,header=TRUE)
```

## 2b
```{r}
coro.fit <- glm(deaths~factor(Age)+factor(Smoke)+factor(Age):factor(Smoke)+offset(log(Pyears)),family=poisson,data=p2)
summary(coro.fit)
```
```{r}
drop1(coro.fit,test = "Chisq")
```
## 2c
Fit the nested version:
```{r}
coro.fit2 <- glm(deaths~factor(Age)+factor(Age):factor(Smoke)+offset(log(Pyears)),family=poisson,data=p2)
summary(coro.fit2)
```
So this data shows a smoke effect in age 1 and age 2.
the mean number of deaths for smoke 1 is exp(1.7469)=5.74 time lager as compared to not-smoke for agegroup 1 and exp(0.7603)=2.14 for agegroup 2.

## 2d
```{r}
pyear.fit <- glm(log(Pyears)~factor(Age)+factor(Smoke),data=p2)
drop1(pyear.fit,test="F")
summary(pyear.fit)
```
There seems to be some logic in these effects.

# Exercise 3

## 3a

```{r}
tick.fit <- glm(TICKS~HEIGHT+YEAR,family=poisson,
                data=lme4::grouseticks)
summary(tick.fit)
```

## 3b
```{r}
 exp(coef(tick.fit))
```
So in 96 the mean number of ticks is 1.51 times larger as in 95 and .185 time lager, so 1/.185 times 5.39 times smaller, in 97 as compared to 95.

## 3c
```{r}
p3 <- data.frame(lme4::grouseticks,res=resid(tick.fit),fv=fitted(tick.fit))
library(ggplot2)
ggplot(p3,aes(x=fv,y=res))+geom_point()
```

There are some fitted values with very large residuals! So the model fitting is not done yet. See dependent data.

# Exercise 4

```{r}
adress <- "https://raw.github.com/janvdbroek/Generalized-Linear-Models/master/BeetleEggCrowding.txt"
p4 <- read.table(adress,header = TRUE)
negg.fit <- glm(NumEggs~factor(Temp)+factor(TRT)+
                  factor(Temp):factor(TRT),
                family = poisson,data=p4)
drop1(negg.fit)
```

Nested version:
```{r}
negg.fit2 <- glm(NumEggs~factor(Temp)+
                  factor(Temp):factor(TRT),
                family = poisson,data=p4)
confint(negg.fit2)
```

So there is a temperature effect at 24 degrees. The log ratio of the mean numer of eggs for trt I vs trt G is -0.5174 with a 95 % profile log-likelihood confidence interval of (-0.827 -0.191).

